### 2020-11-06 latex/acrobot.tex
Finished oscillations and started on rotations

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use \(x_0(t,\eta_0)\) to compute the first-order approximation of
the nonlinear system 
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) satisfies
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    ,
\]
and \(x_1(t,\eta_0)\) is the solution to
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0. \\
    \end{cases}
\end{equation}

We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
The nominal pendulum (\ie when \(I = 0\)) oscillates whenever the
nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
Hence, the domain of oscillations on \(\Gamma\) is given by
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
To find the nominal system we set \(I = 0\);
MATLAB's symbolic toolbox evaluates these expressions as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu} &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\) -- this
is expected, since the nominal system is a single pendulum.
By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).

@@

In the spirit of perturbation analysis, we expand our solution of
(\ref{eqn:muhat-dot}) into
\[
    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
    + R(\alpha,I)
    ,
\]
where (\ref{eqn:khalil-perturbation-nominal}) tells us that
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    .
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
\end{align*}
where we define
\begin{align*}
    K &= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]

\subsection*{Perturbation Analysis for Rotations}

>>>>

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use this to compute the solution \(x_1(t,\eta_0)\) of
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}
Finally, one can compute the first-order approximation
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) is \(O(I^2)\);
that is,
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    .
\]
We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

@@

\subsection*{Perturbation Analysis for Oscillations}
When \(I = 0\), the acrobot becomes a nominal pendulum that oscillates whenever the
nominal energy (\ref{eqn:acrobot-nominal-E}) is less than \(E(\pi,0)\).
Hence, the domain of oscillations on \(\Gamma\) is given by
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
We wish to find a transformation \(T(q_u,p_u)\) on \(D\) into
\((\alpha,\mu)\)-coordinates where \(\alpha \in \mathbb{S}^1\) is a pseudo-angle
and \(\mu > 0\) is a pseudo-radius.
Since the energy of oscillation can be determined by the orbit's intersection
with the \(q_u\)-axis and \(q_u \in ]-\pi,\pi[\) on \(D\), 
we set \(\mu \in ]0,\pi[\) as in Figure \ref{fig:mu-intersection}.
The transformation we want is therefore a diffeomorphism of the form
\begin{align*}
    T : D\backslash \{(0,0)\} &\rightarrow \mathbb{S}^1 \times \, ]0,\pi], \\
    (q_u, p_u) &\mapsto (\alpha,\mu)
    .
\end{align*}

@@

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
The nominal system \((\alpha_n,\mu_n)\) comes from setting \(I = 0\).
MATLAB's symbolic toolbox evaluates the nominal dynamics as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha}_n &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu_n\cos(\alpha_n)) - \cos(\mu_n)}
            {\mu_n^2 \sin(\alpha_n)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha}_n > 0\) for every \(\mu_n \in ]0,\pi[\) -- this
is expected, since the nominal system is a single pendulum.
By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).

@@

In the spirit of perturbation analysis, we expand our solution of
(\ref{eqn:muhat-dot}) into
\begin{equation}\label{eqn:acrobot-muhat-approx}
    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
    + R(\alpha,I)
    ,
\end{equation}
where \(R(\alpha,I)\) is \(O(I^2)\).
From (\ref{eqn:khalil-perturbation-nominal}) we know that 
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    ,
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).

Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]

To show that the first-order approximation (\ref{eqn:acrobot-muhat-approx}) is
gaining energy, we define the Poincar\'{e} section 
\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
in Figure \ref{fig:acrobot-Q} and we see that \(Q(\mu_0)\) is strictly positive,
with an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

\begin{figure}
    \centering
    % TODO: The plot of Q(mu) from MATLAB
    %\includegraphics[width=0.5\textwidth]{images/Qmu.png}
    \caption{The plot of \(Q(\mu_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
Thus, there are no closed orbits and the origin must be a repeller.
Furthermore, for any non-zero initial condition of the acrobot and for small
enough \(I\), the \((q_u,p_u)\) phase of the acrobot will eventually exit \(D\)
and the acrobot will begin rotating.
The acrobot's orbit must leave all compact sets in \(D\) in finite time, which
means the acrobot is gaining energy on \(D\).

If we have \(I < 0\), it is equivalent to using the constraint 
\(q_a = -\bar{q}_a \arctan(I p_u)\). 
By the same analysis, this injects energy on \(D\) in negative-time, so the
constraint dissipates energy on \(D\).

A hybrid supervisor that flips the sign of \(I\) based on the current
value of \(\mu\) will stabilize any energy level of oscillation.
As was the case in the VLP, this supervisor should implement some kind of
hysteresis to avoid infinite switching.
% Empty proof environment to put QED square on the right of the page
\begin{proof}[\unskip\nopunct]
\end{proof}

\subsection*{Perturbation Analysis for Rotations}
Recall that the domain of oscillations for the nominal pendulum of the acrobot
is 
\[
    D := \left\{ (q_u,p_u) \in \Gamma | E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
Outside of \(D\), the acrobot is experiencing rotations.
We wish to find a new set of coordinates \((\alpha,\mu)\) and perform a similar
analysis.
During oscillations we showed that the energy of oscillations is determined by
the intersection with the \(q_u\)-axis; 
when the acrobot is rotating, the energy is instead determined by the
intersection on the \(p_u\)-axis.
Hence, we choose \(\mu\) to be this intersection point.
Note that \(D\) intersects the \(p_u\) axis when 
\(\mu^2 = 60m^2 g l^3\), so we must have 
\(\mu > \sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u > 0\) 
or \(\mu < -\sqrt{60 m^2 g l^3}\) if the rotation intersects at \(p_u < 0\).
The energy level set associated with \(\mu\) is
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\mu^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\mu^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) ranges fully between \(]-pi,\pi]\), so our angle of
rotation is uniquely parameterized by \(\alpha = q_u\).
Since \(\mu\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \alpha
    , \\
    p_u &= \sign{\mu}\sqrt{\mu^2 - 30 m^2 g l^3\left(1 - c_\alpha \right)}
    .
\end{align}

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\), with
inverse 
\begin{align}\label{eqn:rotation-T}
    \alpha &= q_u
    , \\
    \mu &= \sign{p_u}\sqrt{p_u^2 + 30 m^2 g l^3 \left(1 - c_u \right)}
    .
\end{align}

From this point on, we do the exact same procedure as with oscillations.
MATLAB symbolic operations evaluate the nominal system's dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\alpha}_n &=  what
    .
\end{align}
Once again, \(\dot{\alpha}_n > 0\) so for small enough \(I\) we must also have
\(\dot{\alpha} > 0\).
We take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) with initial
\(\mu_0\).
We expand the solution of this into the same form as
(\ref{eqn:acrobot-muhat-approx}) where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \sign{\mu_0}\frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    .
\end{equation}
